Painting with Triangles

### 2014-04-22 paper.tex
figure rearrangement.

\begin{abstract}
Although vector graphics offer a number of benefits, conventional vector painting programs
offer only limited support for the traditional painting metaphor.
We propose a new algorithm that translates a
user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. 
This representation allows the algorithm to render solid colors and linear
gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.

\end{abstract}

@@

\section{Results and Discussion}
Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.  
%
These figures contain drawings in a range of styles. The images shown in \fig{fig:boxes}
reveal a drawing with an extreme zoom range up to 524,000:1, 
for an effective resolution equivalent to an image with more than $10^{17}$ pixels.
At this zoom level our render begins to reach the limits of floating point precision
for single floats; this could be ameliorated by switching to double-precision.

@@

\subsection{Performance}
The current implementation uses Python and PyOpenGL. Python's performance is slower than
a lower level implementation in C++ would be. The drawing of the canvas is quick with
no lag. However, the actual process to take a stroke from rasterization to triangle
representation on the canvas can be slow. On a blank canvas, strokes take less than
one second to process. However, on a canvas with lots of geometry a stroke that overlaps
that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
necessary for the old points affected by the new stroke and the points associated with the
new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
calculations can take an arbitrarily long.
% This brings us to the final point, simplification.

@@

% - Switch 6.3 and 6.4
\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoom}}
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
    \label{fig:face}
\end{figure*}

\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
    \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
    \label{fig:stop}
\end{figure*}

\begin{figure*}
    \centering
        \includegraphics[width=0.195\textwidth]{images/zoom61}
        \includegraphics[width=0.195\textwidth]{images/zoom62}
        \includegraphics[width=0.195\textwidth]{images/zoom63}
        \includegraphics[width=0.195\textwidth]{images/zoom64}
        \includegraphics[width=0.195\textwidth]{images/zoom65}
        \includegraphics[width=0.195\textwidth]{images/zoom66}
        \includegraphics[width=0.195\textwidth]{images/zoom67}
        \includegraphics[width=0.195\textwidth]{images/zoom68}
        \includegraphics[width=0.195\textwidth]{images/zoom69}
        \includegraphics[width=0.195\textwidth]{images/zoom72}
    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1.}
    \label{fig:boxes}
\end{figure*}

\section{Conclusion}
Our system enables a user to paint like they would in a raster graphics program 
to create vector graphics. The core element is a triangle mesh representation of the painting.
We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.

Aside from the simplification challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:

>>>>

\begin{abstract}
Although vector graphics offer a number of benefits, conventional vector painting programs offer only limited support for the traditional painting metaphor. We propose a new algorithm that translates a user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. This representation allows the algorithm to render solid colors and linear gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts. 
\end{abstract}

@@

\section{Results and Discussion}
Paintings made with our system in a range of visual styles are shown in Figures~\ref{fig:teaser} and \ref{fig:face}--\ref{fig:boxes}.  
%
The images shown in \fig{fig:boxes}
reveal a drawing with an extreme zoom range up to 524,000:1, 
for an effective resolution equivalent to an image with more than $10^{17}$ pixels.
At this zoom level our renderer begins to suffer from artifacts at 
the limits of single floating point precision; 
this could be ameliorated by switching to double-precision.

@@

\subsection{Performance}
The current implementation uses Python and PyOpenGL. Python's performance is slower than
a lower level implementation in C++ would be. The preview rasterization of a new stroke on 
the canvas is interactive with
no lag. However, the actual process to take a stroke from rasterization to triangle
representation on the canvas can be slow. On a blank canvas, strokes take less than
one second to process. However, on a canvas with lots of geometry a stroke that overlaps
that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
necessary for the old points affected by the new stroke and the points associated with the
new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
calculations can take an arbitrarily long.
% This brings us to the final point, simplification.

@@

% - Switch 6.3 and 6.4
\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoom}}
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
    \caption{This drawing uses about 70 strokes, resulting in 9030 triangles.}
    \label{fig:face}
\end{figure*}

\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
    \caption{This painting shows soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
    \label{fig:stop}
\end{figure*}

\begin{figure*}
    \centering
        \fbox{\includegraphics[height=0.333\textwidth]{images/apple}}
		\hspace*{0.5in}
        \fbox{\includegraphics[height=0.333\textwidth]{images/eye}}
    \caption{Two more example paintings exploring different styles.}
    \label{fig:apple-eye}
\end{figure*}

\begin{figure*}
    \centering
        \includegraphics[width=0.195\textwidth]{images/zoom61}
        \includegraphics[width=0.195\textwidth]{images/zoom62}
        \includegraphics[width=0.195\textwidth]{images/zoom63}
        \includegraphics[width=0.195\textwidth]{images/zoom64}
        \includegraphics[width=0.195\textwidth]{images/zoom65}
        \includegraphics[width=0.195\textwidth]{images/zoom66}
        \includegraphics[width=0.195\textwidth]{images/zoom67}
        \includegraphics[width=0.195\textwidth]{images/zoom68}
        \includegraphics[width=0.195\textwidth]{images/zoom69}
        \includegraphics[width=0.195\textwidth]{images/zoom72}
    \caption{This single painting shows an extreme zoom; the final closeup image is at zoom level 524,000:1.}
    \label{fig:boxes}
\end{figure*}

\section{Conclusion and Future Work}
Our system enables a user to paint like they would in a raster graphics program 
to create vector graphics. The core element is a triangle mesh representation of the painting.
We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.

Aside from the {\bf simplification} challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:
